---
title: "Aides Drug"
author: "Leonel Bautista"
date: "September 21, 2023"
output: html_document
---

```{r include=FALSE,echo=FALSE}
require(tidyverse)
require(tigerstats)
triUmp <-read.csv(file="triUmpsal.csv",header=TRUE)
```

# Introduction

<!--In this section you explain what you are trying to show.  Where did the data come from?  What is the research or other question you are trying to answer?.  If you skimp in this section expect to be marked down in your grade. Paint a picture of the project a couple of paragraphs is expected!-->

A new  preventative aides drug (triUmpsal) is under test in clinical trials. Consider only the DrugUsers, does this drug seem effective?  Hint - use dplyr to obtain that subpopulation with the filter command.

$H_0$: There is no difference in the data when comparing treatment and infection rate. 

$H_A$: There is a difference in the data when comparing treatment and infection rate. 

```{r include=FALSE}
drug_users <- filter(triUmp, DrugUser == "yes")
```

# Methods

<!--Decide on your methods:  use "variable analysis" or other appropriate descriptors.  Make sure to choose at least one graphical method and at least one numerical method. Discuss what data manipulations may be needed for example subsetting or other binning procedures and what sort of graphical displays and inferential tools will be used!-->

We will use the Dyplr tool "filter" to create a subset of the data, storing only data where drug user= "yes".   

We will then use the CandP helper files to get R code to show a combination of graphical, numerical and inferential to test the null hypothesis. 

# Results

<!--Divide this section into two sub-sections:  One for your descriptive  results and one for your inferential results.!-->

## Descriptive Results

### Graphical Descriptive Results

<!--Graphical results here.  Make sure to show your code.  Provide appropriate labels for axes, giving units if possible, and provide a good title for the graph, too.  Use the graphical results to describe the patterns if any that exist in the data as focused toward the research question. Bottom line is to explain the features of the graph that are relevant to conforming or departing to what is expected by your null hypothesis !-->

```{r}
barchartGC(~infected + treatment,data=drug_users)
barchartGC(~infected + treatment,data=drug_users, type="percent")
```

We have created two bar charts to compare the data, the first shows the total number of people who are involved in the analysis.

In chart 1 we see that most people who are drug users do not contract AIDS, however it does not show us much about the effectiveness of the drug.

In chart 2 we have a clearer picture because it shows the data as a percentage. In this test we see that the treatment group has a lower incidence of AIDS when compared to the placebo group. 

### Numerical Descriptive Results

<!--Numerical results go here. Use the numerical results to describe the patterns if any that exist in the data as focused toward the research question!-->

```{r}
table1 <- xtabs(~infected + treatment, data=drug_users)
rowPerc(table1)
colPerc(table1)
```
In this version of the results we see the number of people in each group and the rate at which they contracted AIDS. 

In the infected group, we see that there are approximately three times as mabt cases of AIDS in the placebo group as opposed to the treatment group.

## Inferential Results

<!--State hypothesis clearly.  Make sure your discussion of the inferential test covers all the aspects that the test output produces, such as test statistic, p-value etc.  Make a decision about the null hypothesis, explain the assumptions on which the selected test/procedure was based, and why the chosen procedure satisfys the assumptions and is appropriate to answer the research question!-->

```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```

We have ran three inferential tests in this section.

Chi-Square Test: This suggests that the observed difference in AIDS contraction rates between the treatment and placebo groups are unlikely to have occurred by chance alone.

Chi-Square Test with Goodness-of-Fit: This supports the notion that there are notable differences in AIDS contraction rates between the treatment and placebo groups.

Fisher's Test: The Fisher's exact test, which is particularly useful for small sample sizes, also yields a significant result (p-value < 0.05). This reinforces the evidence for a non-random association between treatment and AIDS contraction.


# Discussion and Conclusion

<!--Discussion and conclusion here.  If you found a relationship be sure to consider whether the relationship occurs because one of the variavbles causes the other, or whether they perhasps are related for some other reason.  Watch the chapter 6 videos from the GeorgeTown videos collection.!-->

In conclusion, the comprehensive analysis of the data strongly suggests that the "triUmpsal" drug shows promise in reducing the incidence of AIDS among drug users. The graphical, numerical, and inferential results consistently support the notion that the treatment group experiences a lower rate of AIDS contraction compared to the placebo group. These findings provide encouraging evidence for the effectiveness of the drug and warrant further investigation and consideration for broader clinical application.   

The null hypothesis ($H_0$) posited that there would be no difference in the data when comparing treatment and infection rates. The results, however, have led us to decisively reject this hypothesis. Instead, the alternative hypothesis ($H_A$) emerges as the likely scenario: There is indeed a significant difference in AIDS contraction rates between the treatment and placebo groups. This is further corroborated by the statistical tests, including the Chi-Square Test, the Chi-Square Test with Goodness-of-fit, and Fisher's Exact Test, all of which demonstrated a non-random association between treatment and AIDS contraction.   

The study's methodology, which employed a combination of visualizations and rigorous statistical tests, provided a robust framework for drawing meaningful conclusions about the drug's efficacy in this specific population. These findings have the potential to significantly impact the field of AIDS prevention and treatment, offering hope for a more effective and targeted approach in combating the disease within the drug user population. Further research and clinical trials are warranted to validate and expand upon these promising initial results.   
